### 2020-04-18 Paper_EV/Paper_EV.tex
Backup update 18_04_2020

% This is one possible realization of the official FSD Microsoft Word template. The title page has to be filled in in Microsoft Word environment and saved as a PDF file. Note, that in the header and footer the name of the author and the thesis topic have to be provided. 
%
% Compile Template in following order:
%
% pdflatex Paper_EV
% makeindex Paper_EV.nlo -s nomencl.ist -o Paper_EV.nls (Table of Symbols)
% makeindex Paper_EV.acn -s Paper_EV.ist -t Paper_EV.alg -o Paper_EV.acr (Table of Acronyms)
% makeindex Paper_EV.glo -s Paper_EV.ist -t Paper_EV.glg -o Paper_EV.gls
% pdflatex Paper_EV
% pdflatex Paper_EV
%
%
%#################################################################################
%#################################################################################
%#################################################################################

@@

% Set style for pages with arabic numbering 
% This block prints only the chapter number and the chapter name, not the word "Chapter"
%\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\ \thechapter. \ #1}}{}}
% Clear current page style 
\fancyhf{}
% Head (11 pt font size and 12 pt line spacing)
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}
% Foot (11 pt font size and 12 pt line spacing)
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implem. and Eval. of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage \thinspace / \lastpageref{pagesLTS.arabic}}
% Line settings
\renewcommand{\headrulewidth}{0.1pt}	%upper line
\renewcommand{\footrulewidth}{0.1pt}	%lower line

% Set style for pages with roman numbering 
\fancypagestyle{fancy_beginning}{%
%\pagestyle{fancy}
% Clear current page style 
\fancyhf{}%
% Head
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}%
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}%
% Foot
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implem. and Eval. of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}%
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage}%
% Line settings
\renewcommand{\headrulewidth}{0.1pt}%	%upper line
\renewcommand{\footrulewidth}{0.1pt}%	%lower line
}

@@

% Package for acronyms
\usepackage[acronym, automake, nopostdot, nogroupskip]{glossaries}
\setlength{\LTleft}{-0.15cm}			% left justified
\renewcommand*{\arraystretch}{1.4}		% vertical distance between entries, default is 1
% List
\newacronym{EOM}{EOM}{Equations of Motion}
\newacronym{IMU}{IMU}{Inertial Measurement Unit}
\newacronym{INDI}{INDI}{Incremental Nonlinear Dynamic Inversion}
\newacronym{MAV}{MAV}{Micro Air Vehicle}
\newacronym{MCU}{MCU}{Microcontroller}
\newacronym{NED}{NED}{North East Down}
\newacronym{NDI}{NDI}{Nonlinear Dynamic Inversion}
\newacronym{ToF}{ToF}{Time-of-Flight}
% Make gls
\renewcommand*\entryname{Acronym}
\makeglossaries

@@

% For some reason fisrt chapter has a different vert. spacing from the top, so we add additional 0.8cm, starred command makes sure it will not be suppressed at the beginning or at the end of the page
\vspace*{-0.8cm}
\section*{Statutory Declaration}
I, Evghenii Volodscoi, declare on oath towards the Institute of Flight System Dynamics of Technische Universität München, that I have prepared the present Semester Thesis independently and with the aid of nothing but the resources listed in the bibliography.
\\
\\
This thesis has neither as-is nor similarly been submitted to any other university.
\bigskip
\\
\\
Garching,

@@

\section*{Kurzfassung} \label{sec:kurzfassung}
% Start cursive block
\begin{itshape}
Inkrementelle Nichtlineare Dynamische Inversion (INDI) is ein vielversprechender Reglertyp, welcher h\"aufig f\"ur die Regelung von nichtlinearen Flugsystemen unterschiedlicher Art eingesetzt wird. Ein Vorteil der INDI ist, dass dieser Reglertyp kein detalliertes Modell der geregelten Strecke ben\"otigt. Au\ss{}erdem, wirkt INDI sehr effektiv gegen St\"orungen. Diese Semesterarbeit befasst sich mit der Auslegung eines INDI-Reglers f\"ur die Lage- und Positionsregelung eines Nano-Quadrotors. Die Arbeit beginnt mit der Herleitung des Regelalgorithmus. Der hergeleitete Regelgesetz wird dann erst in Simulink entwickelt und anschlie\ss{}end auf der Quadrotor-Hardware implementiert. Dazu werden solche Implementierungsaspekte wie Ermittlung der Eingangsparameter, Messung und Berechnung der Aktuatorzeitkonstante sowie Ermittlung des Schub-Mapping-Parameters besprochen. Abschlie\ss{}end, wird der implementierte Regler auf die F\"ahigkeit St\"orungen zu bek\"ampfen getestet.
\end{itshape}

% Add Abstract chapter to the same page as Kurzfassung
{\let\clearpage\relax\section*{Abstract}}
% Start cursive block
\begin{itshape}
Incremental Nonlinear Dynamic Inversion (INDI) is a promising control technique, widely used for control of different types of aircraft systems. Besides providing high-performance nonlinear control, this controller type does not require a detailed model of the controlled aircraft and is effective against disturbances. This semester thesis describes the development of an INDI controller to control the attitude and the position of a nano-quadrotor. It begins with the derivation of the control algorithm. The controller is then firstly developed in the Simulink environment and afterwards implemented on the quadrotor hardware.  Subsequently, the implementation aspects of the INDI controller such as estimation of the control effectiveness, measurement of the actuator time constant and estimation of the thrust mapping parameter are discussed. Finally, the implemented controller is tested on the ability to cope with disturbances.  
\end{itshape}

@@

The popularity of Micro Air Vehicles (MAVs) in research and industry has grown over the past years. The large interest in \acrshort{MAV}s can be expalined with their potential of performing a wide range of civilian and military tasks, reaching from the investigation of hazardous regions up to the agriculture mapping \cite{Ward}. One of the most popular representatives of \acrshort{MAV}s are quadrotors, belonging to the rotary-wing \acrshort{MAV}s. Their dynamic behaviour is usually of a nonlinear character. Thus, to control such an aircraft, robust control strategies have to be applied. One such strategy is called \acrfull{INDI}, which is the incremental form of the nonlinearity cancelling technique \acrfull{NDI}. The \acrshort{INDI} technique makes it possible to control an aircraft without fully modelling its dynamics. Instead, \acrshort{INDI} uses a primitive aircraft model, which eases the development and implementation of the control algorithm. In addition, \acrshort{INDI} has good properties of coping with unmodelled gust and disturbances.

@@

The main goal of this thesis was to implement an \acrshort{INDI} controller to control the linear acceleration of a nano-quadrotor. Furthermore, to have the ability to control the quadrotors position, the \acrshort{INDI} loop had to be augmented with a velocity as well as a position controller. For the role of the research quadrotor, a small quadrotor named Crazyflie 2.1 was employed. Besides the derivation of the control algorithm itself, this thesis covers the process of estimation of the control effectiveness terms, as well as the estimation of other parameters such as actuator dynamics time constant and thrust mapping parameter, which are necessary for proper functioning of the controller. Additionally to the hardware implementation of the control algorithm, this work describes the development of the controller as well as the quadrotor model in Simulink. Finally, it discusses the capability of the successfully implemented \acrshort{INDI} controller to cope with disturbances.

@@

This semester thesis is organized into five chapters. The introductory chapter is followed by the second chapter, containing the theoretical background. Here, after introducing dynamic equations of motion of an aircraft (\ref{sec:eqs_motion}), \acrshort{NDI} (\ref{sec:ndi}) and \acrshort{INDI} (\ref{sec:indi}) control algorithms are discussed. The \acrshort{INDI} technique is then used to derive the attitude (\ref{subsec: indi_inner}) and position (\ref{subsec: indi_outer}) controllers for the Crazyflie quadrotor. Chapter three focuses on the implementation aspects of the \acrshort{INDI} algorithm derived in the previous chapter. It starts with a short description of the quadrotor hardware (\ref{sec:research_quadrotor}). In the second part (\ref{sec:simulink_model}) of the third chapter, simplified model of the Crazyflie quadrotor as well as the attitude and position controllers are developed in the Simulink environment. The third section (\ref{sec:implementation_on_hardware}) of the same chapter covers aspects that are relevant for the hardware implementation of the \acrshort{INDI} control algorithm, such as the estimation of the control effectiveness terms and thrust mapping parameter. After controller was implemented, it was tested and validated. The test results are summarized in the fourth chapter. The final chapter contains a short conclusion of the performed work.

@@

The aim of this chapter is to provide the theoretical background which serves as a basis for some of the methods which are presented and applied in the course of this thesis. At the beginning, in section \ref{sec:eqs_motion} general \acrfull{EOM} of an aircraft are presented. These are later used to derive the control algorithm and build the Matlab/Simulink model of the Crazyflie quadrotor. Sections \ref{sec:ndi} and \ref{sec:indi} describe the \acrfull{NDI} and \acrfull{INDI} methods in general. Later in this chapter (subsections \ref{subsec: indi_inner} and \ref{subsec: indi_outer}) these methods are used to derive the inner and the outer loop of the \acrshort{INDI} flight controller for the Crazyflie quadrotor.

@@

The Equations (\ref{eq:lin_momentum})-(\ref{eq:position_diff_eq}) presented above can be used to represent a motion of a general aircraft in a three-dimensional space. Note that terms $\sum(\bm{F}^G)_B$ and $\sum (\bm{M}^G)_B$ contain all external forces and moments acting on the rigid body. Considering a general aircraft system, these could be the aerodynamic forces and moments caused by the air flow, propulsion forces and moments, forces caused by the gravitation etc.. The detailed modelling of external forces and moments is presented in chapter \ref{sec:simulink_model}.

@@

Now Equation (\ref{eq:dy_u_rel}) can be used to formulate the control law by solving it for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$
\begin{equation}
	\begin{split}
		\bm{u} = (L_{\bm{G}}\bm{h}(\bm{x}))^{-1} (\bm{\nu} - L_{\bm{f}}\bm{h}(\bm{x}))
		\label{eq:u_control_law}
	\end{split}
\end{equation}
The variable $\bm{\nu}$ is called an \textit{virtual control input} and represents the desired output of the system.

@@

Assuming that the output variable corresponds to the state varible ($\bm{y}=\bm{x}$), the incremental form of the system can be obtained by taking a Taylor series expansion of the Equation \ref{eq:nonlin_sys1}
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{f}(\bm{x}_0) + \bm{G}(\bm{x}_0)\bm{u}_0 \\
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \frac{\partial}{\partial \bm{u}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor}
	\end{split}
\end{equation}
The first term on the right side of the Equation \ref{eq:nonlin_sys_taylor} is $\bm{\dot{x}_0}$. Also, evaluating the differentiation of the third term leads to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{\dot{x}}_0 \\ 
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor2}
	\end{split}
\end{equation}
The second term of the Equation \ref{eq:nonlin_sys_taylor2} contains partial derivative with respect to the state vector. Considering very small time increments of the controller loop and applying the \textit{principle of time scale separation} the second term vanishes. This is a valid assumption if the dynamics of the actuators are fast compared with the dynamics of the system \cite{Silva}. Thus, Equation \ref{eq:nonlin_sys_taylor2} is further simplified to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} = \bm{\dot{x}}_0 + \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_indi}
	\end{split}
\end{equation}
By solving Equation \ref{eq:nonlin_sys_indi} for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$, the \acrshort{INDI} control law is obtained
\begin{equation}
	\begin{split}
		\bm{u} = \bm{u}_0 + \bm{G}(\bm{x}_0)^{-1} (\bm{\nu} - \bm{\dot{x}}_0)
		\label{eq:u_control_law_indi}
	\end{split}
\end{equation}
where $\bm{\dot{x}}_0$ is a measurable value from the previous step, $\bm{u_0}$ the control input from the previous step, $\bm{\nu}$ the reference value and $\bm{G}(\bm{x}_0)$ the control effectiveness matrix. With $\bm{\Delta u}=\bm{u}-\bm{u}_0$ the control law from Equation \ref{eq:u_control_law_indi} represents an incremental version of Equation \ref{eq:u_control_law}. Instead of computing the complete control input command $\bm{u}$, this control law results in computing the increment of the control input $\bm{\Delta u}$ and adding it to the previous value $\bm{u}_0$. As it is less dependent on the model of the system dynamics, it is able to increase the robustness of the system \cite{Sieberling}.

@@

where $\bm{\omega}$ and $\dot{\bm{\omega}}$ are angular velocity and acceleration of the body-fixed frame ($B$) with respect to the \acrshort{NED} ($O$) coordinate frame, $\bm{M}_G$ the gravitational moment, $\bm{M}_A$ the aerodynamic moment and $\bm{M}_P$ the propulsion moment. The vector $\bm{\Omega}$ contains angular velocities of all four rotors and serves as an input variable of the system. It is assumed that the gravitational force is applied to the center of gravity of the quadrotor and does not cause any moment around it. Due to the absence of the aerodynamic moment this term is also omitted and can be seen as a disturbance \cite{Smeur1}. The remaining propulsion moment is written as $\bm{M}_P=\bm{M}_C-\bm{M}_{gyro}$ where $\bm{M}_C$ is the control moment generated by the rotors and $\bm{M}_{gyro}$ the moment containing the gyroscopic effect of the rotors. This two moments can be explicitly written as

@@

where $I_{rzz}$ is the element of the inertia matrix $I_{r}$ of the rotor, $l$ and $b$ lever arms as denoted in the Figure \ref{fig:frames_leverarms}, $k_F$ and $k_M$ force and moment constants of the rotors.

@@

The first term of the Equation \ref{eq:main_eom_indi_inner_lin} is the angular acceleration based on the current angular rates $\bm{\omega}_0$ and inputs $\bm{\Omega}_0$ and can be denoted as $\bm{\dot{\omega}}_0$. $T_0$ is the current thrust value. The last term on the right side of the Equation \ref{eq:main_eom_indi_inner_lin} is scaled with the sample time $T_s$ which is introduced only to simplify further mathematical transformations. The expressions of the control moment $\bm{M}_C$ and the gyroscopic moment of the rotors $\bm{M}_{gyro}$ have been summarized to control effectiveness matrices $\bm{G}_1$ and $\bm{G}_2$

@@

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} laser-ranging sensor which measures the range to the next object along the body-fixed z-axis of the quadcopter. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity of the aircraft.

@@

Before implementing the outer loop of the \acrshort{INDI} controller derived in \ref{subsec: indi_outer} on the real Crazyflie quadrotor it was decided firstly to build a simulation model using Matlab/Simulink software. Such a simulation has a number of advantages compared to the direct implementation on the hardware. The main one is, it allows one to conveniently test different model components (e.g. filtering algorithms, actuator dynamic models etc.) on their dedicated functions, facilitating the debugging process. It also can ease the process of parameter tuning of the controller.

In the framework of this thesis the Simulink model was also applied to better understand the controller structure and its characteristics. Based on the simulation it was determined which of the control effectiveness terms have a greater influence on the system response and which ones can be neglected. Additionally, the simulation model provided a good overview of potential computational problems and mistakes which could lead to drastically growing values and unstable behaviour of the system. To solve some of them, saturation blocks have been used  in different places.

@@

To ensure proper functioning of the inner loop, some of the errors in the already existing implementation 1were corrected. Additionally, the control effectiveness parameters $\bm{G_1}$ and $\bm{G_2}$ were reestimated with a new flight data. For the matrix $\bm{G_1}$ it was assumed that it does not change over the operational domain and can be considered as static.

>>>>

% This is one possible realization of the official FSD Microsoft Word template. The title page has to be filled in in Microsoft Word environment and saved as a PDF file. Note, that in the header and footer the name of the author and the thesis topic have to be provided. 
%
% Compile Template in following order:
%
% pdflatex Paper_EV
% makeindex Paper_EV.nlo -s nomencl.ist -o Paper_EV.nls (Table of Symbols)
% makeindex Paper_EV.acn -s Paper_EV.ist -t Paper_EV.alg -o Paper_EV.acr (Table of Acronyms)
% makeindex Paper_EV.glo -s Paper_EV.ist -t Paper_EV.glg -o Paper_EV.gls
% pdflatex Paper_EV
% pdflatex Paper_EV
%
% TODO:
% - Change date on the title page
% - Change date on the statutory declaration page
%
%#################################################################################
%#################################################################################
%#################################################################################

@@

% Set style for pages with arabic numbering 
% This block prints only the chapter number and the chapter name, not the word "Chapter"
%\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\ \thechapter. \ #1}}{}}
% Clear current page style 
\fancyhf{}
% Head (11 pt font size and 12 pt line spacing)
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}
% Foot (11 pt font size and 12 pt line spacing)
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implementation \& Evaluation of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage \thinspace / \lastpageref{pagesLTS.arabic}}
% Line settings
\renewcommand{\headrulewidth}{0.1pt}	%upper line
\renewcommand{\footrulewidth}{0.1pt}	%lower line

% Set style for pages with roman numbering 
\fancypagestyle{fancy_beginning}{%
%\pagestyle{fancy}
% Clear current page style 
\fancyhf{}%
% Head
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}%
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}%
% Foot
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implementation \& Evaluation of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}%
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage}%
% Line settings
\renewcommand{\headrulewidth}{0.1pt}%	%upper line
\renewcommand{\footrulewidth}{0.1pt}%	%lower line
}

@@

% Package for acronyms
\usepackage[acronym, automake, nopostdot, nogroupskip]{glossaries}
\setlength{\LTleft}{-0.15cm}			% left justified
\renewcommand*{\arraystretch}{1.4}		% vertical distance between entries, default is 1
% List
\newacronym{EOM}{EOM}{Equations of Motion}
\newacronym{IMU}{IMU}{Inertial Measurement Unit}
\newacronym{INDI}{INDI}{Incremental Nonlinear Dynamic Inversion}
\newacronym{IR}{IR}{Infrared}
\newacronym{MAV}{MAV}{Micro Air Vehicle}
\newacronym{MCU}{MCU}{Microcontroller}
\newacronym{NED}{NED}{North East Down}
\newacronym{NDI}{NDI}{Nonlinear Dynamic Inversion}
\newacronym{ToF}{ToF}{Time-of-Flight}
% Make gls
\renewcommand*\entryname{Acronym}
\makeglossaries

@@

% For some reason fisrt chapter has a different vert. spacing from the top, so we add additional 0.8cm, starred command makes sure it will not be suppressed at the beginning or at the end of the page
\vspace*{-0.8cm}
\section*{Statutory Declaration}
I, Evghenii Volodscoi, declare on oath towards the Institute of Flight System Dynamics of Technische Universität München, that I have prepared the present Semester Thesis independently and with the aid of nothing but the resources listed in the bibliography.
\\
\\
This thesis has neither as-is nor similarly been submitted to any other university.
\bigskip
\\
\\
Garching, April 25, 2020
\\
\\
\\
\\
Evghenii Volodscoi

@@

\section*{Kurzfassung} \label{sec:kurzfassung}
% Start cursive block
\begin{itshape}
Inkrementelle Nichtlineare Dynamische Inversion (INDI) is ein vielversprechender Reglertyp, welcher h\"aufig f\"ur die Regelung von nichtlinearen Flugsystemen unterschiedlicher Art eingesetzt wird. Ein Vorteil der INDI ist, dass dieser Reglertyp kein detalliertes Modell der geregelten Strecke ben\"otigt. Au\ss{}erdem, wirkt INDI sehr effektiv gegen St\"orungen. Diese Semesterarbeit befasst sich mit der Auslegung eines INDI-Reglers f\"ur die Lage- und Positionsregelung eines Nano-Quadrotors. Die Arbeit beginnt mit der Herleitung des Regelalgorithmus. Der hergeleitete Regelgesetz wird dann erst in Simulink entwickelt und anschlie\ss{}end auf der Quadrotor-Hardware implementiert. Dazu werden solche Implementierungsaspekte wie Ermittlung der Eingangsmatrix, Messung und Berechnung der Aktuatorzeitkonstante sowie Ermittlung des Schub-Mapping-Parameters besprochen. Abschlie\ss{}end, wird der implementierte Regler auf die F\"ahigkeit St\"orungen zu bek\"ampfen getestet. Der finale Regler ist \"uber das offizielle Open-Source-Firmware des Crazyflie-Quadrotors verf\"ugbar. 
\end{itshape}

% Add Abstract chapter to the same page as Kurzfassung
{\let\clearpage\relax\section*{Abstract}}
% Start cursive block
\begin{itshape}
Incremental Nonlinear Dynamic Inversion (INDI) is a promising control technique, widely used for control of different types of aircraft systems. Besides providing high-performance nonlinear control, this controller type does not require a detailed model of the controlled aircraft and is effective against disturbances. This semester thesis describes the development of an INDI controller to control the attitude and the position of a nano-quadrotor. It begins with the derivation of the control algorithm. The controller is then firstly developed in the Simulink environment and afterwards implemented on the embedded hardware of the quadrotor.  Subsequently, the implementation aspects of the INDI controller such as estimation of the control effectiveness, measurement of the actuator time constant and estimation of the thrust mapping parameter are discussed. Finally, the implemented controller is tested on the ability to cope with disturbance. The final version of the implemented control algorithm is available via official open source firmware of the Crazyflie quadrotor.
\end{itshape}

@@

The popularity of Micro Air Vehicles (MAVs) in research and industry has grown over the past years. The large interest in \acrshort{MAV}s can be explained with their potential of performing a wide range of civilian and military tasks, reaching from the investigation of hazardous regions up to agriculture mapping \cite{Ward}. One of the most popular representatives of \acrshort{MAV}s are quadrotors, belonging to the rotary-wing \acrshort{MAV}s. Their dynamic behaviour is usually of a nonlinear character. Thus, to control such an aircraft, robust control strategies have to be applied. One such strategy is called \acrfull{INDI}, which is the incremental form of the nonlinearity cancelling technique \acrfull{NDI}. The \acrshort{INDI} technique makes it possible to control an aircraft without fully modelling its dynamics. Instead, \acrshort{INDI} uses a primitive aircraft model, which eases the development and implementation of the control algorithm. In addition, \acrshort{INDI} has good properties of coping with unmodelled gust and disturbances.

@@

The main goal of this thesis was to implement an \acrshort{INDI} controller to control the linear acceleration of a nano-quadrotor. Furthermore, to have the ability to control the quadrotors position, the \acrshort{INDI} loop had to be augmented with a velocity as well as a position controller. For the role of the research quadrotor, a small quadrotor named Crazyflie 2.1 was employed. Besides the derivation of the control algorithm itself, this thesis covers the process of estimation of the control effectiveness terms, as well as the estimation of other parameters such as actuator dynamics time constant and thrust mapping parameter, which are necessary for proper functioning of the controller. Additionally to the hardware implementation of the control algorithm, this work describes the development of the controller as well as the quadrotor model in Simulink. Finally, it discusses the capability of the successfully implemented \acrshort{INDI} controller to cope with disturbances. The hardware version of the control algorithm was then made available to general public via official open source firmware.

@@

This semester thesis is organized into five chapters. The introductory chapter is followed by the second chapter, containing the theoretical background. Here, after introducing dynamic equations of motion of an aircraft (\ref{sec:eqs_motion}), \acrshort{NDI} (\ref{sec:ndi}) and \acrshort{INDI} (\ref{sec:indi}) control algorithms are discussed. The \acrshort{INDI} technique is then used to derive the attitude (\ref{subsec: indi_inner}) and position (\ref{subsec: indi_outer}) controllers for the Crazyflie quadrotor. Chapter three focuses on the implementation aspects of the \acrshort{INDI} algorithm derived in the previous chapter. It starts with a short description of the quadrotor hardware (\ref{sec:research_quadrotor}). In the second part (\ref{sec:simulink_model}) of the third chapter, simplified model of the Crazyflie quadrotor as well as the attitude and position controllers are developed in the Simulink environment. The third section (\ref{sec:implementation_on_hardware}) of the same chapter covers aspects that are relevant for the hardware implementation of the \acrshort{INDI} control algorithm, such as the estimation of the control effectiveness terms and thrust mapping parameter. After the controller was implemented, it was tested and validated. The test results are summarized in the fourth chapter. The final chapter contains a short conclusion of the performed work.

@@

The aim of this chapter is to provide the theoretical background which serves as a basis for some of the methods which are presented and applied in the course of this thesis. At the beginning, in section \ref{sec:eqs_motion} general \acrfull{EOM} of an aircraft are presented. These are later used to derive the control algorithm and build the Simulink model of the Crazyflie quadrotor. Sections \ref{sec:ndi} and \ref{sec:indi} describe the \acrfull{NDI} and \acrfull{INDI} methods in general. Later in this chapter (subsections \ref{subsec: indi_inner} and \ref{subsec: indi_outer}) these methods are used to derive the inner and the outer loop of the \acrshort{INDI} flight controller for the Crazyflie quadrotor.

@@

The Equations (\ref{eq:lin_momentum})-(\ref{eq:position_diff_eq}) presented above can be used to represent a motion of a general aircraft in a three-dimensional space. Note that terms $\sum(\bm{F}^G)_B$ and $\sum (\bm{M}^G)_B$ contain all external forces and moments acting on the rigid body. Considering a general aircraft system, these could be the aerodynamic forces and moments caused by the air flow, propulsion forces and moments, forces caused by the gravitation etc.. The detailed modelling of external forces and moments is presented in sections \ref{subsec: indi_inner}, \ref{subsec: indi_outer} and \ref{sec:simulink_model}.

@@

Now Equation (\ref{eq:dy_u_rel}) can be used to formulate the control law by solving it for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$
\begin{equation}
	\begin{split}
		\bm{u} = (L_{\bm{G}}\bm{h}(\bm{x}))^{-1} (\bm{\nu} - L_{\bm{f}}\bm{h}(\bm{x}))
		\label{eq:u_control_law}
	\end{split}
\end{equation}
The variable $\bm{\nu}$ is called a \textit{virtual control input} and represents the desired output of the system.

@@

Assuming that the output variable corresponds to the state varible ($\bm{y}=\bm{h}(\bm{x})=\bm{x}$), the incremental form of the system can be obtained by taking a Taylor series expansion of the Equation \ref{eq:nonlin_sys1}
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{f}(\bm{x}_0) + \bm{G}(\bm{x}_0)\bm{u}_0 \\
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \frac{\partial}{\partial \bm{u}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor}
	\end{split}
\end{equation}
The first term on the right side of the Equation \ref{eq:nonlin_sys_taylor} is $\bm{\dot{x}_0}$. Also, evaluating the differentiation of the third term leads to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{\dot{x}}_0 \\ 
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor2}
	\end{split}
\end{equation}
The second term of the Equation \ref{eq:nonlin_sys_taylor2} contains partial derivative with respect to the state vector. Considering very small time increments of the controller loop and applying the \textit{principle of time scale separation} the second term vanishes. This is a valid assumption if the dynamics of the actuators are fast compared with the dynamics of the system \cite{Silva}. Thus, Equation \ref{eq:nonlin_sys_taylor2} is further simplified to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} = \bm{\dot{x}}_0 + \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_indi}
	\end{split}
\end{equation}
By solving Equation \ref{eq:nonlin_sys_indi} for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$, the \acrshort{INDI} control law is obtained
\begin{equation}
	\begin{split}
		\bm{u} = \bm{u}_0 + \bm{G}(\bm{x}_0)^{-1} (\bm{\nu} - \bm{\dot{x}}_0)
		\label{eq:u_control_law_indi}
	\end{split}
\end{equation}
where $\bm{\dot{x}}_0$ is a measurable value from the previous step, $\bm{u_0}$ the control input from the previous step, $\bm{\nu}$ the reference value and $\bm{G}(\bm{x}_0)$ the control effectiveness matrix. With $\bm{\Delta u}=\bm{u}-\bm{u}_0$ the control law from Equation \ref{eq:u_control_law_indi} represents an incremental version of Equation \ref{eq:u_control_law}. Instead of computing the complete control input command $\bm{u}$, this control law results in computing the increment of the control input $\bm{\Delta u}$ and adding it to the previous value $\bm{u}_0$. As it is less dependent on the model of the system dynamics, it is able to increase the robustness of the system \cite{Sieberling}.

@@

where $\bm{\omega}$ and $\dot{\bm{\omega}}$ are angular velocity and acceleration of the body-fixed frame ($B$) with respect to the \acrshort{NED} ($O$) coordinate frame, $\bm{M}_G$ the gravitational moment, $\bm{M}_A$ the aerodynamic moment and $\bm{M}_P$ the propulsion moment. The vector $\bm{\Omega}$ contains angular velocities of all four rotors and serves as an input variable of the system. It is assumed that the gravitational force is applied to the center of gravity of the quadrotor and does not cause any moment around it. Due to the absence of the aerodynamic moment, this term is also omitted and can be seen as a disturbance \cite{Smeur1}. The remaining propulsion moment is written as $\bm{M}_P=\bm{M}_C-\bm{M}_{gyro}$ where $\bm{M}_C$ is the control moment generated by the rotors and $\bm{M}_{gyro}$ the moment containing the gyroscopic effect of the rotors. This two moments can be explicitly written as

@@

where $I_{rzz}$ is an element of the inertia matrix $I_{r}$ of the rotor, $l$ and $b$ lever arms as denoted in the Figure \ref{fig:frames_leverarms}, $k_F$ and $k_M$ force and moment constants of the rotors.

@@

The first term of the Equation \ref{eq:main_eom_indi_inner_lin} is the angular acceleration based on the current angular rates $\bm{\omega}_0$ and inputs $\bm{\Omega}_0$. $T_0$ is the current thrust value. The last term on the right side of the Equation \ref{eq:main_eom_indi_inner_lin} is scaled with the sample time $T_s$ which is introduced only to simplify further mathematical transformations. The expressions of the control moment $\bm{M}_C$ and the gyroscopic moment of the rotors $\bm{M}_{gyro}$ have been summarized to control effectiveness matrices $\bm{G}_1$ and $\bm{G}_2$ which are defined as

@@

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} laser-ranging sensor which measures the range to the next object along the body-fixed z-axis of the quadrotor. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity of the aircraft.

@@

Before implementing the outer loop of the \acrshort{INDI} controller derived in \ref{subsec: indi_outer} on the real Crazyflie quadrotor it was decided firstly to build a simulation model using the Matlab/Simulink software. Such a simulation has a number of advantages compared to the direct implementation on the hardware. The main one is, it allows one to conveniently test different model components (e.g. filtering algorithms, actuator dynamics models etc.) on their dedicated functions, facilitating the debugging process. It also can ease the process of parameter tuning of the controller.

In the framework of this thesis the Simulink model was also used to better understand the controller structure and its characteristics. Based on the simulation it was determined which of the control effectiveness terms have a greater influence on the system response and which ones can be neglected. Additionally, the simulation model provided a good overview of potential computational problems and mistakes which could lead to drastically growing values and unstable behaviour of the system. To solve some of them, saturation blocks have been used  in different places.

@@

To ensure proper functioning of the inner loop, some of the mistakes in the already existing implementation were corrected. Additionally, the control effectiveness parameters $\bm{G_1}$ and $\bm{G_2}$ were reestimated with a new flight data. For the matrix $\bm{G_1}$ it was assumed that it does not change over the operational domain and can be considered as static.
